---
title: "p8105_hw2_ms6358"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```


## data cleaning

```{r}
litters_data = read_excel("data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx",  "Mr. Trash Wheel", range = "A2:N535") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls))
```

## Read and clean precipitation data for 2018 and 2019

```{r}
precipitation_data_18 = read_excel("data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx","2018 Precipitation", range = "A2:B14")%>%
  janitor::clean_names() %>%
  mutate(year = 2018)


precipitation_data_19 = read_excel("data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx","2019 Precipitation", range = "A2:B14")%>%
  janitor::clean_names() %>%
  mutate(year = 2019)

total_data = 
  bind_rows(precipitation_data_18, precipitation_data_19) %>%
  mutate(month = month.name[month])
```

The number of observations in dataset "litters_data" is `r nrow(litters_data)` and the key variables are `r names(litters_data)`. The number of observations in dataset "total_data" is `r nrow(total_data)` and the key variables are `r names(total_data)`.The total precipitation in 2018 is `r sum(pull(precipitation_data_18, total))`and the median number of sports balls in a dumpster in 2019 is `r median(pull(filter(litters_data, year == 2019), sports_balls))`

## Problem 2

```{r}
pols_df =
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%
  mutate(month = month.name[month]) %>%
  rename(dem = prez_dem) %>%
  rename(gop = prez_gop) %>%
  pivot_longer(c(dem, gop), names_to = "prez", values_to = "president") %>%
  select(-day)
```

```{r}
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day")) %>%
  arrange(year, month)

snp_df_1 = filter(snp_df, year >= 50) %>%
  mutate(year = as.numeric(year) + 1900)

snp_df_2 = filter(snp_df, year <= 15) %>%
  mutate(year = as.numeric(year) + 2000)

snp_df_total = bind_rows(snp_df_1, snp_df_2) %>%
  mutate(month = month.name[as.integer(month)]) %>%
  select(-day) %>%
  relocate(year, month)
```
```{r}
unemployment_df = 
   read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
   pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment rate") %>%
   janitor::clean_names() %>%
   mutate(month = recode(month, `Jan` = "January", `Feb` = "February", `Mar` = "March", `Apr` = "April", `Jun` = "June", `Jul` = "July", `Aug` = "August", `Sep` = "September", `Oct` = "October", `Nov` = "November", `Dec` = "December"))
```

```{r}
polsmonth_snp = left_join(pols_df, snp_df_total)
polsmonth_snp_unemployment = left_join(polsmonth_snp, unemployment_df)
```

The pols-month dataset has dimension `r nrow(pols_df)` x `r ncol(pols_df)`. The  years ranges from 1947 to 2015. The key variables are `r names(pols_df)`. The snp dataset has dimension `r nrow(snp_df_total)` x `r ncol(snp_df_total)`. The years ranges from 1950 to 2015. The key variables are `r names(snp_df_total)`. The unemployment dataset has dimension `r nrow(unemployment_df)` x `r ncol(unemployment_df)`. The years ranges from 1948 to 2015. The key variables are `r names(unemployment_df)`. The joint dataset has dimension `r nrow(polsmonth_snp_unemployment)` x `r ncol(polsmonth_snp_unemployment)`. The  years ranges from 1947 to 2015. The key variables are `r names(polsmonth_snp_unemployment)`.

## Problem 3

```{r}
names_df = read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(ethnicity = recode(ethnicity, `WHITE NON HISP` = "WHITE NON HISPANIC", `ASIAN AND PACI` = "ASIAN AND PACIFIC ISLANDER", `BLACK NON HISP` = "BLACK NON HISPANIC"), childs_first_name = str_to_title(childs_first_name))
```
```{r}
names_filtered = names_df[!duplicated(names_df[,1:6]),]

olivia = filter(names_filtered, childs_first_name == "Olivia", gender == "FEMALE") %>%
  arrange(year_of_birth) %>%
  pivot_wider(ethnicity, names_from = "year_of_birth", values_from = "rank") %>%
  knitr::kable()
olivia

most_popular_male_name = filter(names_filtered, rank == "1", gender == "MALE") %>%
  arrange(year_of_birth) %>%
  pivot_wider(ethnicity, names_from = "year_of_birth", values_from = "childs_first_name") %>%
  knitr::kable()
most_popular_male_name
```
```{r}
male_white_nonhispanic_2016 = filter(names_filtered, gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016) %>% 
  ggplot(aes(x = rank, y = count)) + geom_point()
male_white_nonhispanic_2016
```

  
  
