---
title: "p8105_hw2_ms6358"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```


## data cleaning

```{r}
litters_data = read_excel("data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535") %>%
   janitor::clean_names() %>%
   names() %>%
   drop_na(dumpster) %>%
   mutate(sports_balls = round(sports_balls)) %>%
```

## Read and clean precipitation data for 2018 and 2019

```{r}
precipitation_data_18 = read_excel("data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx","2018 Precipitation", range = "A2:B14")%>%
  janitor::clean_names() %>%
  mutate(year = 2018)


precipitation_data_19 = read_excel("data/Trash-Wheel-Collection-Totals-7-2021-sept.xlsx","2019 Precipitation", range = "A2:B14")%>%
  janitor::clean_names() %>%
  mutate(year = 2019)

total_data = 
  bind_rows(precipitation_data_18, precipitation_2019) %>%
  mutate(month = month.name[month])
```

The number of observations in dataset "litters_data" is `r nrow(litters_data)` and the key variables are `r names(litters_data)`. The number of observations in dataset "total_data" is `r nrow(total_data)` and the key variables are `r names(total_data)`.The total precipitation in 2018 is `r sum(pull(precipitation_data_18, total))`and the median number of sports balls in a dumpster in 2019 is `r median(pull(filter(litters_data, year == 2019). sports_balls))`

## Problem 2

```{r}
pols_df =
  read_csv("./data/__MACOSX/fivethirtyeight_datasets/._pols-month.csv") %>%
  janitor::clean_names() %>%
  seperate(mon, into = c("year", "month", "day")) %>%
  mutate(month = month.name[month]) %>%
  rename(dem = prez_dem) %>%
  rename(gop = prez_gop) %>%
  pivot_longer(c(dem, gop), names_to = "pem", values_to = "president") %>%
  select(-day)
```

```{r}
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  seperate(date, into = c("year", "month", "day")) %>%
  arrange(year, month)
  
  
  
